\bf Calendar-based graphics for visualizing people's daily schedules

### 2019-04-07 reviews/response.Rmd

> Fig.4: Prior to reading the text on p8, I was really confused and even assumed there was a major bug in your R code. Having 2 days in early May and 1 day in early Oct, then 3 days missing, and then the remainder of the month is really confusing for someone who only looks at the figure without reading the text in detail. At least mention this layout anomaly in the figure caption.

Thanks. I have pointed out the layout anomalies for May and October in the Figure 6 caption (previously Figure 4).

> p8, l19: "wrap the last few days up to the top row of the block": This answers
my comment for Fig.4 now, but this layout still remains misleading. We expect to
see some similar temporal pattern in nearby graphs. But there could be considerable differences over a 30-day period, e.g., in your example from Section 3:
What if these 1 or 2 days are after some summer vacation with lots of air conditioning
use, but the vacation already started in the previous month and continued through the middle of the current month. Suddenly, there will be a few huge spikes that interrupt
the overall low-energy pattern. I could think of 2 possible solutions:
(i) Add a 6th week for each month; or (ii) Add the extra days to the
start of the next month. This is sometimes called a
"Calendar Heatmap Tetris Chart", see for example https://stackoverflow.com/questions/27000131/calendar-heat-map-tetris-chart
[both of these features could become additional user options for your R function]

It is intentional to conform a calendar layout of 5 rows and 7 columns for the purpose of compactness. Only one or two months in a year will span over 6 weeks, which would result in more unused space for a year-long calendar plot. We don't shift the extra days to the next month because it is difficult for readers to distinguish which month contains these additional days without proper labelling. But we agree that it is disruptive since time is usually perceived as linear.

That said, the algorithm that lays out the calendar is flexible to incorporate these two options. We have included them as the options for future work. Thanks for the suggestions.

@@

> p11, l44: "star plots": briefly explain what these are and cite a basic reference for these.

Done.

> Fig.7: Where is "noon" and "midnight" here - on top/bottom or on the 0/180 degree position on the right/left side? And what is the direction? I suppose clockwise, but this needs to be mentioned somewhere.

The star plots have been removed from the paper, since it doesn't add more information to other existing plots.

### 2019-04-07 main.Rmd
mentioned suggestions as future work

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \@ref(fig:ped-map)). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

The remainder of the paper is organized as follows. Section \@ref(sec:algorithm) demonstrates the construction of the calendar layout in depth. Section \@ref(sec:transformation) describes the algorithms of data transformation. Section \@ref(sec:opt) lists and describes the options that come with the `frame_calendar` function. Section \@ref(sec:variations) presents some variations of its usage. Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \@ref(sec:case). Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.

@@

The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph]. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month can be captured with 35 (5 $\times$ 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar. Months are of different lengths, ranging from 28 to 31 days, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block. The notation for creating these cells is as follows:

@@

Figure \@ref(fig:month-diagram) illustrates this $(i,j)$ layout for a month where $k=5$.

@@

```{r chn-embedded, fig.cap = "The same plot as Figure \\ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.", eval = FALSE}
include_graphics("figure/chn-1.pdf")
```

@@

@@

The layout is a type of faceting and could be useful to develop this into a fully-fledged faceting method with formal labels and axes. This is a future goal.

>>>>

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \@ref(fig:ped-map)). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. [ToDo: summarise small multiples in more details including citations.] The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

The remainder of the paper is organized as follows. Section \@ref(sec:algorithm) demonstrates the construction of the calendar layout in depth. Section \@ref(sec:transformation) describes the algorithms of data transformation. Section \@ref(sec:opt) lists and describes the options that come with the `frame_calendar` function. Section \@ref(sec:variations) presents some variations of its usage. The calendar layout is a type of faceting, and Section \@ref(sec:facet-calendar) develops this into a fully-fledged faceting method with formal labels and axes. Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \@ref(sec:case). Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.

@@

The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph]. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month conforms to a layout of 5 rows and 7 columns, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar. Months are of different lengths, ranging from 28 to 31 days, and some months could extend over six weeks but for these months we intentionally wrap the last few days up to the top row of the block for the purpose of compact displays. The notation for creating these cells is as follows:

@@

Figure \@ref(fig:month-diagram) illustrates this $(i, j)$ layout for a month where $k=5$.

@@

@@

# A fully-fledged faceting calendar method {#sec:facet-calendar}

@@

### 2019-04-07 main.tex
mentioned suggestions as future work

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia \citep{ped}. There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \ref{fig:ped-map}). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

The remainder of the paper is organized as follows. Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth. Section \ref{sec:transformation} describes the algorithms of data transformation. Section \ref{sec:opt} lists and describes the options that come with the \texttt{frame\_calendar} function. Section \ref{sec:variations} presents some variations of its usage. Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.

@@

The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data \citep{Wickham2012glyph}. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month can be captured with 35 (5 \(\times\) 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar. Months are of different lengths, ranging from 28 to 31 days, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block. The notation for creating these cells is as follows:

@@

Figure \ref{fig:month-diagram} illustrates this \((i,j)\) layout for a month where \(k=5\).

@@

>>>>

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia \citep{ped}. There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \ref{fig:ped-map}). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. {[}ToDo: summarise small multiples in more details including citations.{]} The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

The remainder of the paper is organized as follows. Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth. Section \ref{sec:transformation} describes the algorithms of data transformation. Section \ref{sec:opt} lists and describes the options that come with the \texttt{frame\_calendar} function. Section \ref{sec:variations} presents some variations of its usage. The calendar layout is a type of faceting, and Section \ref{sec:facet-calendar} develops this into a fully-fledged faceting method with formal labels and axes. Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.

@@

The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data \citep{Wickham2012glyph}. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month conforms to a layout of 5 rows and 7 columns, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar. Months are of different lengths, ranging from 28 to 31 days, and some months could extend over six weeks but for these months we intentionally wrap the last few days up to the top row of the block for the purpose of compact displays. The notation for creating these cells is as follows:

@@

Figure \ref{fig:month-diagram} illustrates this \((i, j)\) layout for a month where \(k=5\).

@@

\hypertarget{sec:facet-calendar}{%
\section{A fully-fledged faceting calendar method}\label{sec:facet-calendar}}
